---
title: "Storm Data Analysis"
author: "Zhang Shiwei"
date: "Sunday, September 21, 2014"
output: html_document
---

R Environment Setting
=====================
```{r}
library(knitr)
library(plyr)
library(ggplot2)
Sys.setlocale(category="LC_ALL", locale="English") #this can avoid a strange problem
sessionInfo()
```

Reading and Clearing Data
=========================
```{r}
#download("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2","StormData.csv.bz2")
theData <- read.csv('StormData.csv.bz2', stringsAsFactors=F)
str(theData)
```
I substract some fields that I think useful and droped others.
```{r}
theData <- theData[, c("EVTYPE", "FATALITIES", "INJURIES", "PROPDMG", "PROPDMGEXP", "CROPDMG", "CROPDMGEXP")]
```
Fields "xxEXP" means the exponents, so I use a number to replace them.
```{r}
theData$PROPDMGEXP <- tolower(theData$PROPDMGEXP)
theData$CROPDMGEXP <- tolower(theData$CROPDMGEXP)
replaceEXP <- function(x){
  if(x %in% c("","-","?","+","0")){
    0
  }else if(x == "h"){
    100
  }else if(x == "k"){
    10^3
  }else if(x == "m"){
    10^6
  }else if(x == "b"){
    10^9
  }else{
    10^as.numeric(x)
  }
}
theData$PROPDMG <- theData$PROPDMG*sapply(theData$PROPDMGEXP,replaceEXP)
theData$CROPDMG <- theData$CROPDMG*sapply(theData$CROPDMGEXP,replaceEXP)
theData <- theData[c("EVTYPE", "FATALITIES", "INJURIES", "PROPDMG", "CROPDMG")]
```

Data Analyzing
==============
Figure out influence of each evtype.
```{r}
fatalities <- aggregate(FATALITIES ~ EVTYPE,theData,sum)
injuries <- aggregate(INJURIES ~ EVTYPE,theData,sum)
propdamage <- aggregate(PROPDMG ~ EVTYPE,theData,sum)
cropdamage <- aggregate(CROPDMG ~ EVTYPE,theData,sum)
```
Then sort them.
```{r}
fatalities <- fatalities[order(fatalities$FATALITIES,decreasing=T),]
injuries <- injuries[order(injuries$INJURIES,decreasing=T),]
propdamage <- propdamage[order(propdamage$PROPDMG,decreasing=T),]
cropdamage <- cropdamage[order(cropdamage$CROPDMG,decreasing=T),]
```
OK, now we can plot them and see which influence most.

Result
======
### Top 10 events that are harmful for health
```{r}
par(mfrow=c(1,2),mar=c(8,4,2,2))
barplot(injuries$INJURIES[1:10],names.arg=injuries$EVTYPE[1:10],las=2,main="injuries")
barplot(fatalities$FATALITIES[1:10],names.arg=fatalities$EVTYPE[1:10],las=2,main="fatalities")
```

### Top 10 events that are harmful for economy
```{r}
par(mfrow=c(1,2),mar=c(8,4,2,2))
barplot(cropdamage$CROPDMG[1:10],names.arg=cropdamage$EVTYPE[1:10],las=2,main="crop damage")
barplot(propdamage$PROPDMG[1:10],names.arg=cropdamage$EVTYPE[1:10],las=2,main="prop damage")
```









